Reanalysis Step 2: Multiple Imputation

### 2025-08-26 reanalysis-2-multiple_imputation.qmd
combing through

# ANALYSIS MODE: "test" or "final"
# test = 2 imputations, 5 iterations (fast)
# final = 30 imputations, 20 iterations (slow but thorough)
analysis_mode <- "final"  # Change to "final" for production run

@@

**Default imputation methods:**
```{r method-table}
kable(method_table)

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for ID)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor
custom_methods["ever_lapse_binary"] <- ""  # Complete - use as predictor only
custom_methods["glaucoma_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["otherretina_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["gender_cat"] <- ""  # Likely complete
custom_methods["other_inject"] <- ""  # Likely complete
custom_methods["anti_VEGF"] <- ""  # Likely complete  
custom_methods["focal_laser_flag"] <- ""  # Likely complete
custom_methods["PRP_flag"] <- ""  # Likely complete

# Note: Variables with method = "" (except ID) are STILL USED as predictors for other variables
# They just aren't imputed themselves. The ID is excluded from both imputation and prediction.

@@

# Summary of methods
final_methods <- data.frame(
  Variable = names(custom_methods),
  Method = custom_methods,
  Type = dplyr::case_when(
    custom_methods == "" ~ "Predictor only (not imputed)",
    custom_methods == "pmm" ~ "Predictive mean matching",
    custom_methods == "polr" ~ "Proportional odds regression",
    custom_methods == "logreg" ~ "Logistic regression",
    custom_methods == "polyreg" ~ "Polytomous regression",
    TRUE ~ custom_methods
  ),
  stringsAsFactors = FALSE
)

# Sort to show imputed variables first
final_methods <- final_methods |>
  dplyr::arrange(Method == "", Variable)

vars_to_impute <- names(custom_methods)[custom_methods != ""]
```

**Imputation methods by variable:**
```{r final-methods-table}
kable(final_methods)
```

**Variables that WILL be imputed** (`r length(vars_to_impute)` total):
```{r vars-to-impute}
print(vars_to_impute)

# Show predictor matrix summary
avg_predictors <- sum(pred_matrix[1,])
predictors <- names(which(colSums(pred_matrix) > 0))
```

@@

**Variables used as predictors:**
```{r predictor-list}
print(predictors)
```

>>>>

# ANALYSIS MODE: "test" or "final"
# test = 2 imputations, 5 iterations (fast)
# final = 30 imputations, 20 iterations (slow but thorough)
analysis_mode <- "test"  # Change to "final" for production run

@@

**Initial imputation methods by variable:**
```{r method-table}
kable(method_table)

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for IDs)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor; see below.
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor; see below.

# Variables that are complete (no missing data) - set method but won't actually be imputed
# Setting methods ensures correct measurement level is used IF they had missing data
custom_methods["ever_lapse_binary"] <- "logreg"  # Binary - complete
custom_methods["glaucoma_bef_hitplus_cat"] <- "logreg"  # Binary - already recoded to 0/1
custom_methods["otherretina_bef_hitplus_cat"] <- "logreg"  # Binary - already recoded to 0/1
custom_methods["gender_cat"] <- "logreg"  # Binary categorical - likely complete
custom_methods["other_inject"] <- "logreg"  # Binary - likely complete
custom_methods["anti_VEGF"] <- "logreg"  # Binary - likely complete  
custom_methods["focal_laser_flag"] <- "logreg"  # Binary - likely complete
custom_methods["PRP_flag"] <- "logreg"  # Binary - likely complete

# Note: Variables with complete data won't be imputed even if they have a method assigned
# The method specification ensures proper measurement level if imputation were needed

@@

# Get actual missing data information
missing_info <- data.frame(
  Variable = names(mice_dry$nmis),
  N_Missing = mice_dry$nmis,
  stringsAsFactors = FALSE
)

# Summary of methods
final_methods <- data.frame(
  Variable = names(custom_methods),
  Method = custom_methods,
  Type = dplyr::case_when(
    custom_methods == "" ~ "Excluded (IDs - not used as predictor)",
    custom_methods == "pmm" ~ "Predictive mean matching",
    custom_methods == "polr" ~ "Proportional odds regression",
    custom_methods == "logreg" ~ "Logistic regression",
    custom_methods == "polyreg" ~ "Polytomous regression",
    TRUE ~ custom_methods
  ),
  stringsAsFactors = FALSE
) |>
  left_join(missing_info, by = "Variable") |>
  mutate(
    Will_Be_Imputed = N_Missing > 0 & Method != "",
    Role = case_when(
      Variable %in% c("row_id", "e_mrn_deidentified") ~ "ID (excluded)",
      N_Missing == 0 & Method != "" ~ "Complete predictor",
      N_Missing > 0 & Method != "" ~ "To be imputed",
      TRUE ~ "Other"
    )
  )

# Sort to show meaningful groupings
final_methods <- final_methods |>
  dplyr::arrange(Role == "ID (excluded)", N_Missing == 0, Variable)

# Variables that will ACTUALLY be imputed (have missing data AND a method)
vars_to_impute <- final_methods |>
  filter(Will_Be_Imputed) |>
  pull(Variable)
```

**Imputation methods and missing data by variable:**
```{r final-methods-table}
final_methods |>
  select(Variable, Type, N_Missing, Role) |>
  kable(col.names = c("Variable", "Method Type", "Missing Values", "Role in Imputation"))
```

**Variables that will ACTUALLY be imputed** (have missing data, `r length(vars_to_impute)` total):
```{r vars-to-impute}
if (length(vars_to_impute) > 0) {
  data.frame(
    Variable = vars_to_impute,
    Missing_Count = final_methods |> 
      filter(Variable %in% vars_to_impute) |> 
      pull(N_Missing)
  ) |>
  kable()
} else {
  cat("No variables require imputation (all data complete)")
}

# Show predictor matrix summary
# Calculate average predictors (excluding self-predictions)
avg_predictors <- round(mean(rowSums(pred_matrix[pred_matrix != 0, ])), 1)
predictors <- names(which(colSums(pred_matrix) > 0))

# Verify IDs are excluded
ids_excluded <- all(pred_matrix[, "row_id"] == 0) & all(pred_matrix[, "e_mrn_deidentified"] == 0)
```

@@

**ID variables properly excluded from predictors:** `r if(ids_excluded) "✓ Yes" else "✗ No"`

**Variables used as predictors** (`r length(predictors)` total):
```{r predictor-list}
if ("row_id" %in% predictors || "e_mrn_deidentified" %in% predictors) {
  warning("WARNING: ID variables are being used as predictors! This should not happen.")
}
print(predictors)
```
